---
title: "Prevalencia"
output: html_notebook
---

```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("/Users/danielavargasrobles/Dropbox (UMass Medical School)/Ana_projects/2_PREECLAMPSIA/")) 

```
#LOAD
```{r}

library(phyloseq)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggprism)
library(MASS)
library(vegan)

load("/Users/danielavargasrobles/Dropbox (UMass Medical School)/Ana_projects/2_PREECLAMPSIA/P16S/R/phy_objects/PHYrar_alpha_ggext_subset02_newseqs.RData")
```

# Prevalence tables (no filtering)

```{r}
# --- Inicio del Script ---

physeq <- PHYrar_alpha_ggext_subset02_newseqs

physeq_genus <- tax_glom(physeq, taxrank = "Genus")

# --- Función para Prevalencia y Cuentas ---
calc_prevalence_counts <- function(physeq, group_var = "study_group", taxa_level = "Genus") {
  counts <- as(otu_table(physeq), "matrix")
  sample_inf <- data.frame(sample_data(physeq))
  groups <- sample_inf[[group_var]]
  taxa_names <- as.vector(tax_table(physeq)[, taxa_level])
  taxa_names[is.na(taxa_names)] <- "Unclassified"
  
  results <- data.frame(Genus = taxa_names, stringsAsFactors = FALSE)
  
  for (g in unique(groups)) {
    group_counts <- counts[, groups == g, drop = FALSE]
    prevalence <- rowSums(group_counts > 0) / sum(groups == g) * 100
    count <- rowSums(group_counts > 0)
    results[[paste0("Prevalence_", g)]] <- prevalence
    results[[paste0("Count_", g)]] <- count
  }
  return(results)
}

# --- Función para calcular p-valores Fisher ---
calc_differences_pvalues <- function(prev_table, group1, group2, physeq, group_var = "study_group") {
  results <- prev_table
  sample_metadata <- data.frame(sample_data(physeq))
  total1 <- sum(sample_metadata[[group_var]] == group1)
  total2 <- sum(sample_metadata[[group_var]] == group2)

  results$Difference <- results[[paste0("Prevalence_", group1)]] - results[[paste0("Prevalence_", group2)]]
  
  results$P_value <- sapply(1:nrow(results), function(i) {
    present1 <- results[[paste0("Count_", group1)]][i]
    present2 <- results[[paste0("Count_", group2)]][i]
    absent1 <- total1 - present1
    absent2 <- total2 - present2
    table <- matrix(c(present1, present2, absent1, absent2), nrow = 2, byrow = TRUE)
    fisher.test(table)$p.value
  })
  
  results$Q_value <- p.adjust(results$P_value, method = "fdr")
  return(results)
}

# --- Función para formatear la tabla final ---
format_table <- function(data) {
  data$P_value_formatted <- ifelse(is.na(data$P_value), "NA", sprintf("%.5f", data$P_value))
  data$Significance <- ""
  data$Significance[data$P_value < 0.05] <- "*"
  data$Significance[data$P_value < 0.01] <- "**"
  data$Significance[data$P_value < 0.001] <- "***"
  
  formatted <- data.frame(
    Genus = data$Genus,
    Control_count = data$Count_Control,
    Control_percent = sprintf("%.2f%%", data$Prevalence_Control),
    PE_count = data$Count_PE,
    PE_percent = sprintf("%.2f%%", data$Prevalence_PE),
    Difference = sprintf("%.2f%%", data$Difference),
    P_value = paste0(data$P_value_formatted, data$Significance),
    Q_value = sprintf("%.4f", data$Q_value),
    stringsAsFactors = FALSE
  )
  return(formatted)
}

# --- ANALIZAR POR BODY SITE ---

# Vaginal
physeq_vag <- subset_samples(physeq_genus, body_site == "Vaginal")
prev_vag <- calc_prevalence_counts(physeq_vag)
final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
final_vag <- final_vag[order(final_vag$P_value), ]  # 🔻 ordenar por p-valor
table_vag <- format_table(final_vag)
write.csv(table_vag, "all_indiv_prevalencia_cuentas_Vaginal.csv", row.names = FALSE)

# MomTongue
physeq_mom <- subset_samples(physeq_genus, body_site == "MomTongue")
prev_mom <- calc_prevalence_counts(physeq_mom)
final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
final_mom <- final_mom[order(final_mom$P_value), ]  # 🔻 ordenar por p-valor
table_mom <- format_table(final_mom)
write.csv(table_mom, "all_indiv_prevalencia_cuentas_MomTongue.csv", row.names = FALSE)

# Mostrar resultados ordenados
cat("Tabla para Vaginal (Top 10):\n")
print(head(table_vag, 10))

cat("\nTabla para MomTongue (Top 10):\n")
print(head(table_mom, 10))

```

# Core taxa prevalences (taxa with > 40% oof prevalence) for oral and vaginal
```{r}

# --- 2. Analysis Workflow ---

# Load your initial phyloseq object
physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs

# Agglomerate at Genus level BEFORE filtering
# This is better as prevalence at the genus level is more stable
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
PREVALENCE_THRESHOLD <- 0.40 # 40% prevalence within the body site
ABUNDANCE_THRESHOLD <- 0.00  # 

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply prevalence filter ON THIS SUBSET
  # This creates a logical vector: TRUE for taxa to keep, FALSE for taxa to discard.
  prev_filter <- genefilter_sample(physeq_subset, filterfun_sample(function(x) x > 0), A = PREVALENCE_THRESHOLD * nsamples(physeq_subset))

  # --- FIX IS HERE ---
  # Use the 'prev_filter' variable that you just calculated.
  taxa_to_keep <- prev_filter

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# --- 3. Run Analysis on the independently filtered data ---
# (Vaginal, MomTongue, etc.)


# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    write.csv(table_vag, "P16S/results/prev/site_filtered_prev_40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    write.csv(table_mom, "P16S/results/prev/site_filtered_prev_40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}

```

# Accesory taxa prevalence (taxa with 20-40% of prevalence) for oral and vaginal

```{r}
# Load necessary libraries if you haven't already
# library(phyloseq)
# library(genefilter)

physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs

# Agglomerate at Genus level BEFORE filtering
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
LOWER_PREV_THRESHOLD <- 0.20 # 20% lower prevalence bound
UPPER_PREV_THRESHOLD <- 0.40 # 40% upper prevalence bound
ABUNDANCE_THRESHOLD <- 0.00   # 0% relative abundance threshold

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply filters ON THIS SUBSET

  # Prevalence filter (this part works correctly)
  prevalences <- rowSums(otu_table(physeq_subset) > 0) / nsamples(physeq_subset)
  
  # Example for 20-40% range. You would change this line for other filters.
  taxa_in_range <- prevalences >= 0.20 & prevalences <= 0.40

  # --- FIX IS HERE ---
  # The abundance filter was overriding the prevalence filter.
  # We now use *only* the prevalence filter for pruning.
  taxa_to_keep <- taxa_in_range

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    # Updated filename to reflect the new prevalence range
    write.csv(table_vag, "P16S/results/prev/site_filtered_prev_20-40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    # Updated filename to reflect the new prevalence range
    write.csv(table_mom, "P16S/results/prev/site_filtered_prev_20-40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}
```


# Among White: Core taxa prevalences (taxa with > 40% of prevalence) for oral and vaginal


```{r}
# --- Inicio del Script ---

physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs
physeq_raw=subset_samples(physeq_raw,race == "White/Caucasian" & ethnicity == "Not Hispanic/Latina") 
# tabla prevalancencias con filtro por prev 40

# --- 1. Pre-analysis Functions ---
# NOTA: Las funciones de cálculo (calc_prevalence_counts, etc.) y formato no cambian.
# Eliminamos la función 'filter_taxa_by_prevalence_and_abundance' porque ahora lo haremos de otra forma.

# [Aquí irían las funciones calc_prevalence_counts, calc_differences_pvalues y format_table
# ...son las mismas que en tu script, puedes copiarlas y pegarlas aquí]

# --- 2. Analysis Workflow ---



# Agglomerate at Genus level BEFORE filtering
# This is better as prevalence at the genus level is more stable
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
PREVALENCE_THRESHOLD <- 0.40 # 40% prevalence within the body site
ABUNDANCE_THRESHOLD <- 0.00  # 

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply prevalence filter ON THIS SUBSET
  # This creates a logical vector: TRUE for taxa to keep, FALSE for taxa to discard.
  prev_filter <- genefilter_sample(physeq_subset, filterfun_sample(function(x) x > 0), A = PREVALENCE_THRESHOLD * nsamples(physeq_subset))

  # --- FIX IS HERE ---
  # Use the 'prev_filter' variable that you just calculated.
  taxa_to_keep <- prev_filter

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# El resto del script para analizar y escribir los resultados es correcto...
# --- 3. Run Analysis on the independently filtered data ---
# (Vaginal, MomTongue, etc.)


# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    write.csv(table_vag, "P16S/results/prev/white_site_filtered_prev_40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    write.csv(table_mom, "P16S/results/prev/white_site_filtered_prev_40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}



```
# Among Hispanics: Accesory taxa prevalence (taxa with 20-40% of prevalence) for oral and vaginal


```{r}
# Load necessary libraries if you haven't already
# library(phyloseq)
# library(genefilter)

physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs
physeq_raw <- subset_samples(physeq_raw,   ethnicity == "Hispanic/Latina")

# Agglomerate at Genus level BEFORE filtering
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
LOWER_PREV_THRESHOLD <- 0.20 # 20% lower prevalence bound
UPPER_PREV_THRESHOLD <- 0.40 # 40% upper prevalence bound
ABUNDANCE_THRESHOLD <- 0.00   # 0% relative abundance threshold

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply filters ON THIS SUBSET

  # Prevalence filter (this part works correctly)
  prevalences <- rowSums(otu_table(physeq_subset) > 0) / nsamples(physeq_subset)
  
  # Example for 20-40% range. You would change this line for other filters.
  taxa_in_range <- prevalences >= 0.20 & prevalences <= 0.40

  # --- FIX IS HERE ---
  # The abundance filter was overriding the prevalence filter.
  # We now use *only* the prevalence filter for pruning.
  taxa_to_keep <- taxa_in_range

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    # Updated filename to reflect the new prevalence range
    write.csv(table_vag, "P16S/results/prev/hispa_site_filtered_prev_20-40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    # Updated filename to reflect the new prevalence range
    write.csv(table_mom, "P16S/results/prev/hispa_site_filtered_prev_20-40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}
```

# Among White: Accesory taxa prevalence (taxa with 20-40% of prevalence) for oral and vaginal

```{r}

physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs
#physeq=subset_samples(physeq,race == "White/Caucasian" & ethnicity == "Not Hispanic/Latina") 

physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs
physeq_raw=subset_samples(physeq_raw,race == "White/Caucasian" & ethnicity == "Not Hispanic/Latina") # tabla prevalancencias con filtro por prev 20-40

# Load necessary libraries if you haven't already
# library(phyloseq)
# library(genefilter)


# Agglomerate at Genus level BEFORE filtering
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
LOWER_PREV_THRESHOLD <- 0.20 # 20% lower prevalence bound
UPPER_PREV_THRESHOLD <- 0.40 # 40% upper prevalence bound
ABUNDANCE_THRESHOLD <- 0.00   # 0% relative abundance threshold

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply filters ON THIS SUBSET

  # Prevalence filter (this part works correctly)
  prevalences <- rowSums(otu_table(physeq_subset) > 0) / nsamples(physeq_subset)
  
  # Example for 20-40% range. You would change this line for other filters.
  taxa_in_range <- prevalences >= 0.20 & prevalences <= 0.40

  # --- FIX IS HERE ---
  # The abundance filter was overriding the prevalence filter.
  # We now use *only* the prevalence filter for pruning.
  taxa_to_keep <- taxa_in_range

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    # Updated filename to reflect the new prevalence range
    write.csv(table_vag, "P16S/results/prev/white_site_filtered_prev_20-40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    # Updated filename to reflect the new prevalence range
    write.csv(table_mom, "P16S/results/prev/white_site_filtered_prev_20-40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}
```

# Among Hispanics: Core taxa prevalences (taxa with > 40% of prevalence) for oral and vaginal

```{r}
# --- 1. Pre-analysis Functions ---
# NOTA: Las funciones de cálculo (calc_prevalence_counts, etc.) y formato no cambian.
# Eliminamos la función 'filter_taxa_by_prevalence_and_abundance' porque ahora lo haremos de otra forma.

# [Aquí irían las funciones calc_prevalence_counts, calc_differences_pvalues y format_table
# ...son las mismas que en tu script, puedes copiarlas y pegarlas aquí]

# --- 2. Analysis Workflow ---

# Load your initial phyloseq object
physeq_raw <- PHYrar_alpha_ggext_subset02_newseqs
physeq_raw <- subset_samples(physeq_raw,   ethnicity == "Hispanic/Latina")
# Agglomerate at Genus level BEFORE filtering
# This is better as prevalence at the genus level is more stable
physeq_genus_raw <- tax_glom(physeq_raw, taxrank = "Genus", NArm = FALSE)

# --- Define filtering criteria ---
PREVALENCE_THRESHOLD <- 0.40 # 40% prevalence within the body site
ABUNDANCE_THRESHOLD <- 0.00  # 

# --- Create a list to store filtered data for each site ---
physeq_site_filtered_list <- list()

# Get all unique body sites from your data
body_sites <- unique(as.character(sample_data(physeq_genus_raw)$body_site))

# --- Loop through each body site to filter it independently ---
for (site in body_sites) {
  cat(paste("\n--- Filtering site:", site, "---\n"))

  # 1. Subset the data for the current site
  physeq_subset <- subset_samples(physeq_genus_raw, body_site == site)

  # 2. Apply prevalence filter ON THIS SUBSET
  # This creates a logical vector: TRUE for taxa to keep, FALSE for taxa to discard.
  prev_filter <- genefilter_sample(physeq_subset, filterfun_sample(function(x) x > 0), A = PREVALENCE_THRESHOLD * nsamples(physeq_subset))

  # --- FIX IS HERE ---
  # Use the 'prev_filter' variable that you just calculated.
  taxa_to_keep <- prev_filter

  # 3. Prune taxa for the current site
  physeq_site_filtered <- prune_taxa(taxa_to_keep, physeq_subset)

  cat(paste("Original taxa in this site:", ntaxa(physeq_subset), "\n"))
  cat(paste("Taxa remaining after filtering:", ntaxa(physeq_site_filtered), "\n"))

  # Store the filtered phyloseq object in the list
  physeq_site_filtered_list[[site]] <- physeq_site_filtered
}

# El resto del script para analizar y escribir los resultados es correcto...
# --- 3. Run Analysis on the independently filtered data ---
# (Vaginal, MomTongue, etc.)


# --- 3. Run Analysis on the independently filtered data ---

# Vaginal
physeq_vag <- physeq_site_filtered_list[["Vaginal"]]
if (!is.null(physeq_vag) && nsamples(physeq_vag) > 0 && ntaxa(physeq_vag) > 0) {
    prev_vag <- calc_prevalence_counts(physeq_vag)
    final_vag <- calc_differences_pvalues(prev_vag, "PE", "Control", physeq_vag)
    final_vag <- final_vag[order(final_vag$P_value), ]
    table_vag <- format_table(final_vag)
    write.csv(table_vag, "P16S/results/prev/hispa_site_filtered_prev_40_Vaginal.csv", row.names = FALSE)
    cat("\nResults for Vaginal:\n")
    print(table_vag)
} else {
    cat("\nNo data left for Vaginal site after filtering.\n")
}

# MomTongue
physeq_mom <- physeq_site_filtered_list[["MomTongue"]]
if (!is.null(physeq_mom) && nsamples(physeq_mom) > 0 && ntaxa(physeq_mom) > 0) {
    prev_mom <- calc_prevalence_counts(physeq_mom)
    final_mom <- calc_differences_pvalues(prev_mom, "PE", "Control", physeq_mom)
    final_mom <- final_mom[order(final_mom$P_value), ]
    table_mom <- format_table(final_mom)
    write.csv(table_mom, "P16S/results/prev/hispa_site_filtered_prev_40_MomTongue.csv", row.names = FALSE)
    cat("\nResults for MomTongue:\n")
    print(table_mom)
} else {
    cat("\nNo data left for MomTongue site after filtering.\n")
}

```